---
title: "R Notebook"
output: html_notebook
---

FINEMAP

http://www.christianbenner.com/
https://www.ncbi.nlm.nih.gov/pmc/articles/PMC4495769/

Instructions: https://variani.github.io/finemapr/vignettes/finemapr.html

If using straight FINEMAP, need to include inputs:
Master file 
Z file - minor allele frequencies, beta, and se in z-file
LD file
BGEN, BGI, SAMPLE, and INCL file

https://github.com/topics/finemap
https://github.com/variani/finemapr and https://variani.github.io/finemapr/vignettes/finemapr.html
https://github.com/slwest/finemap-scripts

Google search: FINEMAP
https://github.com/Ensembl/Finemap
https://github.com/jinghuazhao/FM-pipeline
https://github.com/manning-lab/fineMap
https://github.com/chr1swallace/finemap-psa
https://github.com/opentargets/genetics-finemapping
https://omictools.com/finemap-tool

1000 Genomes
https://www.google.com/search?q=background+finemap&oq=background+finemap&aqs=chrome..69i57.1595j0j1&sourceid=chrome&ie=UTF-8
https://www.semanticscholar.org/paper/FINEMAP%3A-efficient-variable-selection-using-summary-Benner-Spencer/08f23c6b03aa51eb83948b750162b63810f838cc
https://www.google.com/search?q=1000+genomes+plink+files&oq=1000+genomes+&aqs=chrome.4.69i57j0l5.4991j0j1&sourceid=chrome&ie=UTF-8
https://www.google.com/search?q=1000+genomes+background+score+regression&oq=1000+genomes+background+score+regression&aqs=chrome..69i57.3654j0j1&sourceid=chrome&ie=UTF-8
http://www.internationalgenome.org/data/

```{r load_packages}

# Load finemapr
library(devtools)
devtools::load_all("/Users/kathryntsai/Documents/GitHub/finemapr")
#devtools::install_github("variani/finemapr")

library(magrittr)
library(dplyr)
library(ggplot2)
library(knitr)
theme_set(theme_linedraw())

```

All three fine-mapping tools estimated the poterior causal probabilities of the two variants, rs15 and rs47, very close to 1.

```{r example_data_load}

file1_z <- system.file("extdata/region1.z", package = "finemapr")
file1_ld <- system.file("extdata/region1.ld", package = "finemapr")
z1 <- read_zscore(file1_z)
ld1 <- read_ld(file1_ld, snps = z1$snp)
n1 <- 5363

```

```{r explore z_scores}

z1 %>% arrange(-abs(zscore)) %>% head(5) %>% kable(digits = 1)
ggplot(z1, aes(zscore)) + geom_histogram()
mutate(z1, pval = pchisq(zscore^2, df = 1, lower.tail = FALSE)) %>%
  ggplot(aes(pval)) + geom_histogram()
```

```{r run_finemap}

options(finemapr_finemap = "/Users/kathryntsai/finemap_v1.3.1_MacOSX/finemap_v1.3.1_MacOSX")
out_finemap <- run_finemap(z1, ld1, n1, args = "--n-causal-max 3")
print(out_finemap)
plot(out_finemap, label_size = 3, grid_ncol = 1)

```

./CAVIAR 
-z /Users/kathryntsai/OneDrive\ -\ Villanova\ University/College/2018-2019/Summer\ 2019/TFs_eQTLs_Research/RProjects/Project2/finemapping/Zscore_file_ACBD6.Z 
-l  /Users/kathryntsai/OneDrive\ -\ Villanova\ University/College/2018-2019/Summer\ 2019/TFs_eQTLs_Research/RProjects/Project2/finemapping/ld_matrix_ACBD6.ld 
-c 5 
-o /Users/kathryntsai/OneDrive\ -\ Villanova\ University/College/2018-2019/Summer\ 2019/TFs_eQTLs_Research/RProjects/Project2/finemapping/max_casual5 
-f 1 
-r 0.95

```{r run_caviar}

options(finemapr_caviar = "~/apps/caviar/CAVIAR")
out_caviar <- run_caviar(z1, ld1, args = "-c 3")
print(out_caviar)
plot(out_caviar, label_size = 3)

```

```{r run_paintor}

options(finemapr_paintor = "~/apps/paintor/PAINTOR")
out_paintor <- run_paintor(z1, ld1, n1, args = "-enumerate 3")
print(out_paintor)
plot(out_paintor, label_size = 3)

```

