---
title: "Make a Plink file"
output: html_notebook
---

https://github.com/bulik/ldsc

Todo:
- make plink files with each cis eQTL, organized by chromosome
- run plink in system on each
- if the matrix is too slow, take pairwise snp comparisons and make matrix in r

```{r install_libraries}

#source("https://bioconductor.org/biocLite.R")
#biocLite("snpStats") # https://www.bioconductor.org/packages//2.10/bioc/html/snpStats.html

# library(snpStats)
# library(BiocManager)
# library(biomaRt)
# library(snpEnrichment) # https://rdrr.io/cran/snpEnrichment/man/writeLD.html

```

```{r specify_data}

# http://webcache.googleusercontent.com/search?q=cache:http://zzz.bwh.harvard.edu/plink/dosage.shtml#format

chr1_cis <- franke_cis_data[which(GeneChr==1),]
chr1_cis_split <- split(chr1_cis, chr1_cis$GeneSymbol)
ACBD6 <- chr1_cis_split[[11]] # ACBD6
```

# FOR PLINK FILES

```{r create_snp_file_1_gene_locus}

fwrite(data.table(ACBD6$SNP), "/Users/kathryntsai/OneDrive\ -\ Villanova\ University/College/2018-2019/Summer\ 2019/TFs_eQTLs_Research/RProjects/Project2/finemapping/acbd6_snps.txt", row.names=F, col.names=F)

```

```{r make_map_1_gene_locus}

map_file <- data.table(ACBD6$SNPChr, ACBD6$SNP, 0, ACBD6$SNPPos)
fwrite(map_file, "/Users/kathryntsai/OneDrive\ -\ Villanova\ University/College/2018-2019/Summer\ 2019/TFs_eQTLs_Research/RProjects/Project2/finemapping/acbd6.map", sep="\t", col.names = F, row.names=F)
```

https://web.archive.org/web/20190528204434/http://zzz.bwh.harvard.edu/plink/data.shtml
Have to make up information for summary stats:
     Family ID
     Individual ID
     Paternal ID
     Maternal ID
     Sex (1=male; 2=female; other=unknown)
     Phenotype
http://cubre.covenantuniversity.edu.ng/wp-content/uploads/2014/07/H3ABionet_2014_GWAS_2_Plink_Data_Format.pdf

```{r make_ped_1_gene_locus, include=F}

ACBD6_table <- data.table(ACBD6$SNP, ACBD6$OtherAllele, ACBD6$AssessedAllele)
colnames(ACBD6_table) <- c("SNP", "Ref", "Alt")
#ACBD6_table_melted <- melt(ACBD6_table, id.vars="SNP") 
#ACBD6_table_melted_alleles <- ACBD6_table_melted[order(ACBD6_table_melted$SNP), "value"]

#cbind(ACBD6_table[seq(1, nrow(ACBD6_table), by=2), -1], ACBD6_table[seq(2, nrow(ACBD6_table), by=2), -1])

acbd6_alleles <- strsplit(paste0(paste0(ACBD6_table$Ref, ACBD6_table$Alt), sep="", collapse=""), split="")

ped_file <- data.table("chr1", "acbd6", 0, 0, 3, -9, t(acbd6_alleles[[1]]))
fwrite(ped_file, "/Users/kathryntsai/OneDrive\ -\ Villanova\ University/College/2018-2019/Summer\ 2019/TFs_eQTLs_Research/RProjects/Project2/finemapping/acbd6.ped", sep="\t", col.names = F, row.names=F)
```

```{r plink_test, include=F}

# in plink_mac_20190907 folder
# create test.ped, test.map --> create toy_analysis.frq, toy_analysis.log
system("/Users/kathryntsai/plink-1.07-mac-intel/plink --file test --freq --out toy_analysis --noweb") # getting started
# test.fam, test.bim, test.bed, test.log
system("/Users/kathryntsai/plink-1.07-mac-intel/plink --file test --maf 0.05 --make-bed --out test") # https://www.cog-genomics.org/plink/1.9/data#make_bed
# plink.ld, plink.log
system("/Users/kathryntsai/plink-1.07-mac-intel/plink --file test --r")
```

https://www.cog-genomics.org/plink/1.9/ld 
https://www.cog-genomics.org/plink/1.9/data#make_bed
https://web.archive.org/web/20181008145054/http://zzz.bwh.harvard.edu/plink/ld.shtml

```{r run_plink_part1_for_1_gene_locus}

# in /Users/kathryntsai/OneDrive - Villanova University/College/2018-2019/Summer 2019/TFs_eQTLs_Research/RProjects/Project2/finemapping
system("/Users/kathryntsai/plink-1.07-mac-intel/plink --file acbd6 --allow-no-sex --make-bed --out acbd6") 

```

```{r run_plink_part2_unused_for_1_gene_locus, include=F}

system("/Users/kathryntsai/plink-1.07-mac-intel/plink --file acbd6 --r") # pairwise snps

system("/Users/kathryntsai/plink-1.07-mac-intel/plink --r2 --bfile acbd6 --extract acbd6_snps.txt --ld-window-r2 0.8 --out acbd6_ld")

# https://www.cog-genomics.org/plink2/resources
# https://bioinformatics.stackexchange.com/questions/2905/using-plink-to-find-snps-in-ld-linkage-disequilibrium-with-another-set-of-snps

system("/Users/kathryntsai/plink-1.07-mac-intel/plink --r2 bin --bfile 1kg_phase1_chr1 --extract acbd6_snps.txt --ld-window-r2 0.8 --out acbd6_ld")

# https://bioinformatics.stackexchange.com/questions/2905/using-plink-to-find-snps-in-ld-linkage-disequilibrium-with-another-set-of-snps
# https://www.cog-genomics.org/plink2/resources
system("/Users/kathryntsai/plink-1.07-mac-intel/plink --r2 --bfile 1kg_phase1_chr1 --clump acbd6_snps.txt --clump-field P --clump-p1 1.0001 --clump-p2 0.0001 --clump-r2 0.8 --clump-kb 1000 --out acbd6_ld3")

# error
system("/Users/kathryntsai/plink-1.07-mac-intel/plink --bfile 1kg_phase1_all --r2 --ld-window-r2 0 --out acbd6_ld4")

system("/Users/kathryntsai/plink-1.07-mac-intel/plink --bfile 1kg_phase1_chr1 --r2 --matrix --out acbd6_ld5")

```

https://www.biostars.org/p/263373/
https://www.biostars.org/p/274562/

https://www.bioconductor.org/packages//2.10/bioc/manuals/snpStats/man/snpStats.pdf

# SNPSTATS

```{r snpStats, include=F}

data(testdata)
ld1 <- ld(Autosomes[, 1:50], depth=10, stats=c("D.prime", "R.squared"))[["R.squared"]]
ld2 <- ld(Autosomes[, 1:50], depth=9, stats=c("D.prime", "R.squared"))
ld3 <- ld(Autosomes[, 1:20], Autosomes[, 21:25], stats="R.squared") 

```

# PROXYSNPS

```{r proxysnps, include=F}

d <- get_proxies(chrom = "1", pos = 180448558, window_size = 665391, pop = "AFR")

```

# LDMATRIX

no column names

```{r ldmatrix_false, include=F}
fwrite(data.table(glue_collapse(ACBD6$SNP, sep="\\n")), "acbd6_snps_collapsed.txt")

system('curl -k -H "Content-Type: application/json" -X POST -d "{"snps": "acbd6_snps_collapsed.txt", "pop": "CEU","r2_d": "r2"}" "https://ldlink.nci.nih.gov/LDlinkRest/ldmatrix?&token=bc2543d7f1e5" > "/Users/kathryntsai/OneDrive\ -\ Villanova\ University/College/2018-2019/Summer\ 2019/TFs_eQTLs_Research/RProjects/Project2/finemapping/abcd6_ld_main_ceu.txt"')


```

column names

### curl -k -H "Content-Type: application/json" -X POST -d '{"snps": "chr1:181058272\nrs10157891\nrs10158367\nrs1043744\nrs10465516\nrs10494527\nrs10733004\nrs10733005\nrs10737342\nrs10737343\nrs10737344\nrs10737345\nrs10752833\nrs10753208\nrs10753209\nrs10753210\nrs10753211\nrs10753212\nrs10753213\nrs10753214\nrs10753215\nrs10753216\nrs10753217\nrs10753218\nrs10798744\nrs10798745\nrs10798746\nrs10798749\nrs10798752\nrs10798753\nrs10798754\nrs10798756\nrs10798757\nrs10798759\nrs10798760\nrs10798761\nrs10798762\nrs10798763\nrs10798764\nrs10798766\nrs10798767\nrs10798769\nrs10798770\nrs10798771\nrs10798772\nrs10798775\nrs10913968\nrs10913972\nrs10913973\nrs10913974\nrs10913976\nrs10913978\nrs10913980\nrs10913981\nrs10913982\nrs10913983\nrs10913984\nrs10913985\nrs10913986\nrs10913987\nrs10914001\nrs10914002\nrs10914005\nrs10914010\nrs10914011\nrs10914014\nrs10914016\nrs10914017\nrs10914018\nrs10914019\nrs10914020\nrs10914022\nrs10914023\nrs10914028\nrs10914041\nrs10914042\nrs10914043\nrs10914046\nrs10914047\nrs10914128\nrs1094597\nrs1094598\nrs1094599\nrs113448263\nrs114459820\nrs114517401\nrs115179436\nrs115570388\nrs115663406\nrs11580793\nrs115844726\nrs11586077\nrs116412143\nrs116420224\nrs116451573\nrs11799870\nrs11806607\nrs119465\nrs12024201\nrs12027553\nrs12032249\nrs12034217\nrs12045636\nrs12077904\nrs12117593\nrs12118635\nrs12121043\nrs12122338\nrs12122997\nrs12123254\nrs12127886\nrs12128560\nrs12129511\nrs12129784\nrs12131071\nrs12131915\nrs12132837\nrs12135463\nrs12136111\nrs12141348\nrs12141848\nrs12142124\nrs12143350\nrs12144149\nrs12238995\nrs12240121\nrs12407684\nrs12561776\nrs12565152\nrs12565492\nrs12691484\nrs12730889\nrs12733492\nrs12749611\nrs1339722\nrs1339723\nrs1339726\nrs1361363\nrs1361628\nrs139573699\nrs141389405\nrs1416821\nrs1416823\nrs1418100\nrs1418102\nrs1418103\nrs142452264\nrs146187127\nrs146286448\nrs147052174\nrs149537486\nrs151231956\nrs151335857\nrs1538620\nrs1538622\nrs1556935\nrs1556938\nrs1591898\nrs16856127\nrs17300950\nrs180872319\nrs187483246\nrs188017296\nrs1925339\nrs192878539\nrs1934410\nrs1934411\nrs1935283\nrs1961741\nrs1970401\nrs2105232\nrs2147774\nrs2153327\nrs2153753\nrs2262371\nrs2296378\nrs2331545\nrs2331739\nrs2331741\nrs2331742\nrs2331743\nrs2764432\nrs2764434\nrs2764435\nrs2764436\nrs2764437\nrs2764438\nrs2764439\nrs2764440\nrs2764441\nrs2764442\nrs2764444\nrs2764448\nrs2764449\nrs2764451\nrs2764452\nrs2764453\nrs2764454\nrs2764455\nrs2764456\nrs2764457\nrs2764458\nrs2764459\nrs2764461\nrs2764462\nrs2764463\nrs2764464\nrs2764465\nrs2764467\nrs2764468\nrs2794960\nrs2794962\nrs2794963\nrs2794964\nrs2794965\nrs2794966\nrs2794967\nrs2794968\nrs2794969\nrs2794970\nrs2794971\nrs2794972\nrs2794973\nrs2794974\nrs2794975\nrs2794976\nrs2794978\nrs2877399\nrs34569745\nrs357045\nrs357049\nrs357050\nrs3845397\nrs3856071\nrs3923222\nrs4058058\nrs4076403\nrs4083947\nrs4111061\nrs4142992\nrs4259605\nrs4281291\nrs4304539\nrs4333819\nrs4418555\nrs4420052\nrs4421553\nrs4424469\nrs4436345\nrs4442335\nrs4454504\nrs4483365\nrs4517313\nrs4532796\nrs4540612\nrs4562579\nrs4578179\nrs4593778\nrs4607831\nrs4609392\nrs4634864\nrs4642852\nrs4651061\nrs4651062\nrs4651063\nrs4651064\nrs4651066\nrs4651067\nrs4652493\nrs4652495\nrs4652496\nrs4652498\nrs4652499\nrs4652501\nrs4652502\nrs4652504\nrs4652505\nrs4652506\nrs4652507\nrs4652510\nrs55740741\nrs55874160\nrs56300828\nrs58276994\nrs59019561\nrs61811560\nrs61811627\nrs61811628\nrs61811629\nrs61811630\nrs61811631\nrs6425618\nrs6425620\nrs6425621\nrs6425622\nrs6425623\nrs6425624\nrs6425625\nrs6425626\nrs6425630\nrs6425642\nrs6425646\nrs6425647\nrs6657492\nrs6659147\nrs6659513\nrs6661208\nrs6664756\nrs6664944\nrs6665615\nrs6665719\nrs6667478\nrs6669304\nrs6670544\nrs6671901\nrs6674244\nrs6677367\nrs6678216\nrs6678261\nrs6678305\nrs6680361\nrs6680368\nrs6680754\nrs6682067\nrs6683484\nrs6684842\nrs6687873\nrs6688844\nrs6691333\nrs6694877\nrs6695360\nrs6697924\nrs6700986\nrs6704010\nrs697261\nrs697262\nrs704334\nrs704336\nrs716760\nrs72712997\nrs74134326\nrs74889023\nrs74928106\nrs7512134\nrs7515052\nrs75173421\nrs75190206\nrs7522928\nrs7523003\nrs7523857\nrs7526743\nrs7527034\nrs7528650\nrs7529042\nrs7530015\nrs7532806\nrs7532972\nrs7536561\nrs7538219\nrs7538303\nrs7539910\nrs7540276\nrs75406886\nrs7541051\nrs7542168\nrs7542568\nrs7543143\nrs7543366\nrs7544585\nrs7546030\nrs75468761\nrs7548707\nrs7548819\nrs7548884\nrs7555087\nrs75852301\nrs75898898\nrs76044604\nrs76223943\nrs76238232\nrs76284034\nrs76463494\nrs76937829\nrs77026599\nrs77389071\nrs77561757\nrs77626476\nrs77653878\nrs78013293\nrs78079295\nrs78111936\nrs78142091\nrs78256914\nrs78561363\nrs78639761\nrs78676246\nrs79120763\nrs79418492\nrs79680073\nrs79762611\nrs80205419\nrs80261041\nrs822703\nrs822704\nrs822706\nrs822724\nrs863720\nrs863721\nrs878641\nrs880911\nrs912766\nrs9425497\nrs9425879\nrs946633\nrs946634\nrs946635\nrs9729819\nrs9943120", "pop": "CEU","r2_d": "r2"}' 'https://ldlink.nci.nih.gov/LDlinkRest/ldmatrix?&token=bc2543d7f1e5' > '/Users/kathryntsai/OneDrive - Villanova University/College/2018-2019/Summer 2019/TFs_eQTLs_Research/RProjects/Project2/finemapping/abcd6_ld.ld'

```{r ldmatrix}

ldin <- glue_collapse(ACBD6$SNP, sep="\\n")
sysin <- noquote(gsub("\\n", "\n", gsub('\\\'', "\'", paste('curl -k -H "Content-Type: application/json" -X POST -d \'{', '"snps"', ": \"", ldin, '", "pop": "CEU","r2_d": "r2"}\'', " 'https://ldlink.nci.nih.gov/LDlinkRest/ldmatrix?&token=bc2543d7f1e5' > '/Users/kathryntsai/OneDrive\ -\ Villanova\ University/College/2018-2019/Summer\ 2019/TFs_eQTLs_Research/RProjects/Project2/finemapping/abcd6_ld.ld'", sep=""), str, fixed=T), str, fixed=T))

# doesn't work but idk why --> just use manual

z <- system2(sysin) 

```

